---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(dxhelper)
data("coding_dict")
data("cohort")
data("data_dict")
```

# dxhelper

<!-- badges: start -->
<!-- badges: end -->

The goal of dxhelper is to provide an tidy interface to decoding raw values from exported Apollo datasets, whether they are exported from `dx extract_dataset` or through a SparkSQL query.

Given the data dictionary and the codings files, it will decode the categorical values from the raw data.

## Installation

You can install the development version of dxhelper from [GitHub](https://github.com/) with:

``` r
install.packages("remotes")
remotes::install_github("laderast/dxr")
```

## Note about example data

All examples are shown with synthetic data. These are not actual patients from UKB RAP.

## Decoding Integer Categories of Apollo Datasets

Categorical data is returned by `dx extract_dataset` as the integer representation. To see the actual values, they must be decoded from the `codings.csv` that is generated from `dx extract_dataset`. 

Data fields that have multiple categories must also be parsed and decoded. If the data is a sparse field, it is a combined categorical/numerical field.

## Build coding dictionary

We'll first build a coding dictionary by combining `coding_dict` and `data_dict`:

```{r}
merged_code <- merge_coding_data_dict(coding_dict, data_dict)

head(merged_code)
```

## Show Raw Cohort File

Here's a cohort that was returned by `dx extract_dataset`. Note that there are multiple columns encoded as categories. We need to decode these values.

```{r}
knitr::kable(head(cohort[,c(1:5,14)]))
```



## Decode Columns

We can decode the integer values into the actual values using `decode_categories` and our `merged_code` object:

```{r}
decoded_cohort <- cohort |>
  decode_df(merged_code)

knitr::kable(head(decoded_cohort[,c(1:5,14)]))
```

## Decode Column Names

Optionally, you can also decode the column names in the table to be R-friendly column names with the actual field titles:

```{r}
final_cohort <- decoded_cohort |>
  decode_column_names(merged_code)

knitr::kable(head(final_cohort[,c(1:5,14)]))
```

## Multi-valued columns

Multivalued columns are returned as a comma delimited string:

```{r}
knitr::kable(final_cohort[1:10,"diagnoses_main_icd10"])
```
