---
title: "Running Table Exporter"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(xvhelper)
```

## Running Table Exporter

If we have a large number of fields (more than 15-20) to extract from the pheno data, then our call to `extract_data()` may fail. That is because this functionality is dependent on a shared resource called the Thrift Server. There is a hard limit to the query execution time on the Thrift server: 2 minutes.

If our query takes longer, we can launch Table Exporter, which is an app on the platform that will do the extraction for us. This vignette outlines how to launch table exporter in your R session, monitor it, and find the CSV file that was generated by it.

The first thing we need to do is find our dataset's id, and have a vector of fields that we've generated. Once we have these two items, we can use `launch_table_exporter()` to start the Table Exporter app. 

```{r}
ds_id <- find_dataset_id()
fields <- c("participant.eid", "participant.p31", "participant.p41202")
job_id <- launch_table_exporter(ds_id, fields)
job_id
```

When our Table Exporter job is running, we can check on its status using `check_job()`:

```{r}
check_job(job_id)
```

Note that it also returns a `NULL`. When our job finishes successfully, it will return a `file-id` (see below).

The states our job can be are:

- `idle` 
- `runnable`
- `running`
- `failed`
- `done`

If we need to terminate our Table Exporter Job, we can use `terminate_job()`:

```{r}
terminate_job(job_id)
```

## Successful Table Exporter Run

If our job finishes successfully or fails, we will receive an email notifiation. If we check on our job again with `check_job()`. Here we're passing in a job ID for a successful run.

```{r}
file_id <- check_job("job-GY4Zj180Yq3BJyFzg2ygGVX2")
file_id
```

## Finding all jobs

We can see a list of all jobs and all their states by using `find_all_jobs()`

```{r}
job_frame <- find_all_jobs()
job_frame
```

We can find the finished jobs by looking for `state == "done"`:

```{r}
job_frame |>
  dplyr::filter(state == "done")
```
